---
title: "fit-multiple-models"
author: "Rick O. Gilmore"
date: "`r Sys.Date()`"
output: 
  html_document:
    self_contained: false
    code_folding: show
    toc: true
    toc_depth: 2
    toc_float: true
  github_document: default
params:
  SubID: 1059
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include=FALSE)

# load libraries
require(dplyr)
require(ggplot2)
require(knitr)
require(psyphy) # for mafc
require(lme4)

source("analyses/normalize.child.data.R")
```

## Estimate probit

```{r full-probit-model}
form.full <- Acc ~ Coh + Speed + PatternType + Coh:Speed + Coh:PatternType + (1|SubID)
mod.probit.full <- glmer(formula = form.full, family=binomial(mafc.probit(2)), data = df)
```

```{r summary-probit-model, include=TRUE}
summary(mod.probit.full)
```

Add probit fit to data frame.

```{r add-probit}
df$Probit <- predict(mod.probit.full, type="response")
```

## Estimate logit

```{r full-logit-model}
mod.logit.full <- update(mod.probit.full, family=binomial(mafc.logit(2)))
```

```{r summary-logit-model, include=TRUE}
summary(mod.logit.full)
```

Add fit to data frame.

```{r add-logit}
df$Logit <- predict(mod.logit.full, type="response")
```

## Weibull fit

```{r full-weibull-model}
mod.weib.full <- update(mod.probit.full, family=binomial(mafc.weib(2)))
```

```{r summary-weibull-model, include=TRUE}
summary(mod.weib.full)
```

Add fit to data frame.

```{r add-weibull}
df$Weibull <- predict(mod.weib.full, type="response")
```

## Linear

```{r full-linear-model}
mod.linear.full <- lmer(formula = form.full, data = df)
```

```{r summary-weibull-model, include=TRUE}
summary(mod.linear.full)
```

## Compare models

### probit v. logit 

```{r probit-v-logit}
anova(mod.probit.full, mod.logit.full)
```

Logit is better fit.

### logit vs. Weibull

```{r probit-v-logit}
anova(mod.logit.full, mod.weib.full)
```

Logit is the better fit over the Weibull.

## Logit vs. linear

```{r probit-v-logit}
anova(mod.logit.full, mod.linear.full)
```

Linear model is the best fit!

## Fine-tuning linear model

It appears we can drop speed.

```{r drop-speed}
form.nospeed <- Acc ~ Coh + PatternType + Coh:Speed + Coh:PatternType + (1|SubID)
mod.linear.nospeed <- update(mod.linear.full, formula=form.nospeed)
anova(mod.linear.full, mod.linear.nospeed)
```
This doesn't change things much. Let's also drop the Coh:Speed interaction.

```{r drop-speed}
form.nospeed.2 <- Acc ~ Coh + PatternType + Coh:PatternType + (1|SubID)
mod.linear.nospeed.2 <- update(mod.linear.full, formula=form.nospeed.2)
anova(mod.linear.full, mod.linear.nospeed.2)
```

This suggests it's wise to keep the full model.

## Plotting observed vs. fitted

Trim the data frame to essential variables, including the linear fits.

```{r trim-df-to-plot}
df$Pred <- predict(mod.linear.full, type="response")
df %>% 
  group_by(AgeYrs, Gender, SubID, PatternType, Speed, Coh, Pred) %>% 
  summarize(N.corr = sum(Acc), 
            N.tot = n(), 
            Pct.Corr = N.corr/N.tot,
            RT.mean=mean(RT),
            RT.sd=sd(RT)) -> 
  df.small
```

### Observed

```{r p-corr-plot}
df.small %>%
  ggplot() +
  aes(x=Coh, y=Pct.Corr) +
  geom_point() +
  facet_grid(facets = Speed ~ PatternType)
```

### Fitted

```{r p-corr-plot}
df.small %>%
  ggplot() +
  aes(x=Coh, y=Pred) +
  geom_point() +
  facet_grid(facets = Speed ~ PatternType)
```

## Observed vs. fitted

```{r observed-v-fitted}
df.small %>%
  ggplot() +
  aes(x=Pct.Corr, y=Pred) +
  geom_point() +
  facet_grid(facets = Speed ~ PatternType)
```
