---
title: "model-fitting-children"
author: "Rick O. Gilmore"
date: "`r Sys.Date()`"
output: 
  html_document:
    self_contained: false
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: true
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include=FALSE)

# load libraries
require(dplyr)
require(ggplot2)
require(knitr)
require(psyphy) # for mafc
require(lme4)

source("analyses/normalize.child.data.R")
```

## Estimate full probit model

```{r full.model}
form.full <- Acc ~ Coh + Speed + PatternType + Coh:Speed + Coh:PatternType + (1|SubID)
mod.full <- glmer(formula = form.full, family=binomial(mafc.probit(2)), data = df)
```

```{r summary.full.model, include=TRUE}
summary(mod.full)
```

## Estimate reduced model

Drop Coh:PatternType from model.

```{r drop.coh.patt.interaction}
form.reduced.1 <- Acc ~ Coh + Speed + PatternType + Coh:Speed + (1|SubID)
mod.reduced.1 <- update(mod.full, formula = form.reduced.1)
```

```{r summary.drop.coh.patt.interaction, include=TRUE}
summary(mod.reduced.1)
```

Now, we compare this reduced model to the full model.

```{r drop.vs.full, include = TRUE}
anova(mod.full, mod.reduced.1)
```

This improves the fit. We observe that this model says *p*(correct) increases with increasing coherence, is larger for radial patterns, and the slope of the coherence by *p*(correct) relation is steeper at the faster (8 deg/s) speed.

## Testing age effect

Our plots suggest that there may be an effect of age in years. So, we estimate the influence of age in years as a fixed effect.

```{r age.yrs}
form.age <- Acc ~ Coh + Speed + PatternType + AgeYrs + Coh:Speed + (1|SubID)
mod.age <- update(mod.full, formula = form.age)
```

```{r summary.mod.age}
summary(mod.age)
anova(mod.reduced.1, mod.age)
```

The model including age in years has a better fit. We note that the *p*(corr) is larger for each of the age groups, and that there is an ordinal relationship to the age-related influences on *p*(corr).

## Plots

### *p*(corr) by coherence

We choose the `mod.reduced.1` from above to plot.

For plotting purposes, it's easier to generate a new `df.small` data frame.

```{r add-pred-to-df}
df$Pct.Corr.Pred <- predict(mod.reduced.1, type="response")
df %>% 
  group_by(AgeYrs, Gender, SubID, PatternType, Speed, Coh, Pct.Corr.Pred) %>% 
  summarize(N.corr = sum(Acc), 
            N.tot = n(), 
            Pct.Corr = N.corr/N.tot,
            RT.mean=mean(RT),
            RT.sd=sd(RT)) -> 
  df.small
```

```{r p-corr-plot}
df.small %>%
  filter(SubID == 1059) %>%
  ggplot() +
  aes(x=Coh, y=Pct.Corr) +
  geom_point() +
  geom_point(aes(y=Pct.Corr.Pred, color = "blue")) +
  facet_grid(facets = Speed ~ PatternType)
```

## Resources

### R

```{r session-info}
sessionInfo()
```